---
title: "Gap analysis"
output:
  pdf_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
  html_document:
    toc: true
---
```{r databasetableanalysed}

message("A lot of the code below has to be rewritten because
        missing data is not zero anymore. 
        Missing data is now correctly encoded 
        as NULL in the database 
        and imported as NA values in R.")

# Comment out to pass tableanalysed as a parameter
tableanalysed = 'raw_comext_monthly_201709' 

# Render this long document in a separate bash console with the command:
# cd ~/R/eutradeflows && Rscript -e "tableanalysed = 'raw_comext_monthly_201709'; rmarkdown::render('docs/gapanalysis.Rmd')"
```


```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir="..") # file paths are relative to the root of the project directory
library(eutradeflows)
library(dplyr)
library(dbplyr)
library(tidyr)
library(ggplot2)

# Use rotated text at some point to place country names in table heading
# https://tex.stackexchange.com/questions/201696/how-to-rotate-text-in-a-cell-and-center-it-vertically-and-horizontally
``` 


Connect to the database.
```{r dbConnect}
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
```


# Introduction
Following analysis is based on the most recent monthly data downloaded from the [Eurostat Comext bulk repository](http://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&dir=comext).


## Summary
The eurostat Comext database contains information on import and export flows
between european countries and over 200 partner countries. 
Each flow between a reporter and a partner country contains 3 informations: trade value, weight and quantity.
All 3 informations given by the Comext trade database can have gaps.
Gaps are measured below in terms of number of database rows for which a given value is equal to zero.
There are far fewer gaps in trade value compared to weight and quantity.
```{sql comparegaps, connection = con, tab.cap="Gaps in the 3 variables present in Comext"}
select * from 
(select count(*) as 'gaps in trade value' from raw_comext_monthly_201709 where tradevalue = 0 ) as x, 
(select count(*) as 'gaps in weight' from raw_comext_monthly_201709 where weight = 0) as y, 
(select count(*) as 'gaps in quantity' from raw_comext_monthly_201709 where quantity = 0) as z, 
(select count(*) as 'total number of rows' from raw_comext_monthly_201709) as t;
```


## Descriptive statistics of the recent Comext monthly data
Descriptive queries on the database table raw_comext_monthly.
Number of rows in the table:
```{sql rowsintable, connection=con}
select count(*) from raw_comext_monthly_201709;
```

Number of products
```{sql countproduct, connection=con}
select count(distinct(productcode)) from raw_comext_monthly_201709;
```

Number of reporter and partner countries
```{sql countreporter, connection=con}
select count(distinct(reportercode)) as 'number of reporters',
       count(distinct(partnercode))  as 'number of partners' 
from raw_comext_monthly_201709;
```

First and last month
```{sql minperiod, connection=con}
select min(period), max(period) from raw_comext_monthly_201709;
```


# Gaps in trade value
Number of database rows for which the trade value is equal to 0. 
```{r gapsintradevalue, eval=FALSE}
gtv <- tbl(con, tableanalysed) %>% 
    filter(tradevalue == 0) %>% 
    group_by(productcode, reportercode, period) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by="reportercode") %>% 
    collect() %>% ungroup() %>%  
    mutate(productcode = as.character(productcode),
           prod2 = as.numeric(substr(productcode, 1,2)))
gtv %>%
    group_by(reporter, prod2) %>% 
    summarise(n = sum(n)) %>% 
    left_join(tradeharvester::products2harvest, 
              by = c("prod2" = "productcode")) %>% 
    unite(product, prod2, productdescription) %>% 
    spread(product, n, fill="") %>% 
    kable()

gtv %>%
    mutate(year = substr(as.character(period),1,4)) %>% 
    group_by(reporter, year) %>% 
    summarise(n = sum(n)) %>% 
    spread(year, n, fill="") %>% 
    kable()
```


Heat map plots inspired by a blog post on 
[learnr.wordpress.com](https://learnr.wordpress.com/2010/01/26/ggplot2-quick-heatmap-plotting/).

```{r gtv_period,fig.cap="Number of gaps in trade value by country and period", fig.height=10}
gtv_period <- tbl(con, tableanalysed) %>% 
    filter(tradevalue == 0) %>% 
    group_by(reportercode, period) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by = "reportercode") %>% 
    collect() %>% 
    mutate(period = as.character(period))

ggplot(gtv_period, aes(reporter, period)) +
    geom_tile(aes(fill = n), colour = "white") + 
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


```{r gtv_product,fig.cap="Number of gaps in trade value by country and product", fig.height=10}
gtv_product <- tbl(con, tableanalysed) %>% 
    filter(tradevalue == 0) %>% 
    group_by(reportercode, productcode) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by = "reportercode") %>% 
    collect() %>% 
    mutate(productcode = as.character(productcode),
           p4 = substr(productcode, 1,4)) 

ggplot(gtv_product, aes(reporter, p4)) +
    geom_tile(aes(fill = n), colour = "white") + 
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```
There was a time in 2015 where Croatia had a great number of gaps
in the trade value, but in 2017, this has been mostly fixed.


# Gaps in weight data

Number of lines for which the weight is equal to zero.

```{r gapsinweight, eval=FALSE}
gw <- tbl(con, tableanalysed) %>% 
    filter(weight == 0) %>% 
    group_by(productcode, reportercode, period) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by="reportercode") %>% 
    collect() %>% ungroup() %>%  
    mutate(productcode = as.character(productcode),
           prod2 = as.numeric(substr(productcode, 1,2)))
gw %>%
    group_by(reporter, prod2) %>% 
    summarise(n = sum(n)) %>% 
    left_join(tradeharvester::products2harvest, 
              by = c("prod2" = "productcode")) %>% 
    unite(product, prod2, productdescription) %>% 
    spread(product, n, fill="") %>% 
    kable()

gw %>%
    mutate(year = substr(as.character(period),1,4)) %>% 
    group_by(reporter, year) %>% 
    summarise(n = sum(n)) %>% 
    spread(year, n, fill="") %>% 
    kable()
```




```{r fig.cap="Number of gaps in weight data by country and period", fig.height=10}
gw_period <- tbl(con, tableanalysed) %>% 
    filter(weight == 0) %>% 
    group_by(reportercode, period) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by = "reportercode") %>% 
    collect() %>% 
    mutate(period = as.character(period))

ggplot(gw_period, aes(reporter, period)) +
    geom_tile(aes(fill = n), colour = "white") + 
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r fig.cap="Number of gaps in weight data by country and product", fig.height=10}
gw_product <- tbl(con, tableanalysed) %>% 
    filter(weight== 0) %>% 
    group_by(reportercode, productcode) %>% 
    summarise(n = n()) %>% 
    left_join(tbl(con, "vld_comext_reporter"), by = "reportercode") %>% 
    collect() %>% 
    mutate(productcode = as.character(productcode),
           p4 = substr(productcode, 1,4)) 

ggplot(gw_product, aes(reporter, p4)) +
    geom_tile(aes(fill = n), colour = "white") + 
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Gaps in volume data

# Combine gaps
## Gaps in weight *or* volume data
```{r}
# Number of rows which have missing weight and quantity
dtf <- tbl(con, tableanalysed) %>% 
    group_by(productcode, reportercode, flowcode) %>% 
    mutate(weightna = is.na(weight),
           weight0 = weight==0,
           quantityna = is.na(quantity),
           quantity0 = quantity == 0) %>% 
    summarise(nrow = n(), # total number of rows
              weightna = sum(weightna),
              weight0 = sum(weight0),
              quantityna = sum(quantityna),
              quantity0 = sum(quantity0)) %>% 
    collect() %>% 
    # keep only products at the 8 digit level
    filter(nchar(productcode) == 8)

# Missing data by great product code
dtf %>% 
    mutate(productcode2d = substr(productcode, 1,2)) %>% 
    group_by(productcode2d) %>% 
    summarise(nrow = sum(nrow),
            weightna = sum(weightna),
            weight0 = sum(weight0, na.rm=TRUE),
            quantityna = sum(quantityna),
            quantity0 = sum(quantity0, na.rm = TRUE))
dtf %>% 
    mutate(productcode2d = substr(productcode, 1,2)) %>% 
    group_by(productcode2d) %>% 
    summarise(weightna_percent = round(sum(weightna) 
                                       /sum(nrow) * 100,1),
              quantityna_percent = round(sum(quantityna) /
                                             sum(nrow)*100,1)) %>% 
    kable()

# Missing data by reporter for product code 44
dtf %>% 
    filter(substr(productcode, 1,2) == "44") %>% 
    group_by(reportercode) %>% 
    summarise(weightna_percent = round(sum(weightna) 
                                       /sum(nrow) * 100,1),
              quantityna_percent = round(sum(quantityna) /
                                             sum(nrow)*100,1)) %>% 
    kable()
```

## Gaps in weight where quantity data is there
```{r}
dtf <- tbl(con, tableanalysed) %>% 
    # filter(is.na(weight) & !is.na(quantity)) %>% 
    group_by(productcode, reportercode, flowcode) %>% 
    mutate(weightna = is.na(weight),
           quantityna = is.na(quantity),
           weightnaquantitynona = is.na(weight) & !is.na(quantity)) %>% 
    summarise(nrow = n(), # total number of rows
              weightna = sum(weightna),
              quantityna = sum(quantityna),
              weightnaquantitynona = sum(weightnaquantitynona)
              ) %>% 
    collect() %>% 
    # keep only products at the 8 digit level
    filter(nchar(productcode) == 8)

dtf %>% 
    mutate(productcode2d = substr(productcode, 1,2)) %>% 
    group_by(productcode2d) %>% 
    summarise(weightnaquantitynona_percent = round(sum(weightnaquantitynona) 
                                       /sum(nrow) * 100,1)) %>% 
    kable()
```



# Issues
## 20170925 Issue with the encoding of empty values mysql NULL, R NA
How are empty values encoded in the text file?
```{bash eval=FALSE}
cp ~/R/tradeharvester/data-raw/comext/201709/data/nc201501.7z /tmp/comext
cd /tmp/comext/
7zr e nc201501.7z 
# Extract all productcode starting with 44
awk -F, '$3 ~ /^44/' nc201501.dat | less
```

In `less`, search for /2675.752 
Empty value is represented by an empty string at the end of the line.
Compared for example the end of the 2 lines below:
```
001,0053,44,1,4,201501,2675.752,3247.776,
001,0053,44032011,2,4,201501,35.48,391.5,870
```

Can this be read as NA instead of 0?
How about in the middle of the line?
It seems empty values are then encoded as "0" when in the middle of the line:
```
003,0960,44,1,4,201501,0.157,0,
003,0960,44,2,4,201501,0.141,0.003,
003,0960,44190090,2,4,201501,0.139,0.003,
003,0960,44MMM000,1,4,201501,0.157,0,
003,0960,44MMM000,2,4,201501,0.002,0,
```

# Product descriptions

```{r}
tbl(con, tableanalysed) %>% 
    distinct(productcode) %>% 
    left_join(tbl(con, "vld_comext_product"), by = "productcode") %>% 
    collect() %>% 
    # correct encoding issue, see docs/productcodes.Rmd
    mutate(productdescription = gsub(", MA.*ARANDUBA",", MACARANDUBA",
                                     productdescription)) %>% 
    kable()
```



Disconnect from the database.
```{r dbDisconnect}
RMariaDB::dbDisconnect(con)
```

