---
title: "Create a database to contain Comext data"
author: "Paul Rougieux"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Create a database to contain Comext data}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
output:
  knitr:::html_vignette:
    toc: yes
---


```{r buildvignette, eval=FALSE, echo=FALSE}
# This vignette should be edited in the /vignettes folder
# Documentation on how to build vignettes in http://r-pkgs.had.co.nz/vignettes.html
# To build this vignette run 
devtools::build_vignettes()
devtools::build(vignettes = TRUE)
# Then rebuild the package CTRL+SHIFT+B so that the vignette is 
# integrated to the package documentation
# Maybe also run check?
```


```{r setup, warning=FALSE, message=FALSE}
library(knitr)
# Do not evaluate code chunks below, only display code
opts_chunk$set(eval=FALSE) 
library(dplyr)
```


## Install a MariaDB database server and client

Install mariadb server and client on Debian
```
sudo apt installmariadb-server
sudo apt install libmariadbclient-dev
```

Make sure the tradeflows package is also installed. 
It contains the database configuration in inst/config/databaseconfig.R
Note: this should probably be changed to environment variables. 


## Create a database

In bash enter the following commands:
```
sudo mariadb
```
Then at the MariaDB prompt:
```
CREATE USER 'R'@'localhost' IDENTIFIED BY 'localhost';
CREATE USER paul@localhost;
create database tradeflows;
connect tradeflows;
GRANT ALL PRIVILEGES ON tradeflows. * TO R@localhost;
GRANT ALL PRIVILEGES ON tradeflows. * TO paul@localhost;
quit
```

## Create a database structure
Recreating the empty database structure:

* will not erase the main  raw and validated data. 
 Because only templates are created. 
* it will erase all complementary raw tables raw_comext_partner, raw_comext_product, raw_comext_reporter, raw_comext_unit, raw_comext_unit_description
* and it will erase all complementary validated tables vld_comext_partner, vld_comext_price, vld_comext_priceconversion , vld_comext_pricew , vld_comext_product, vld_comext_reporter, vld_comext_unit, vld_comext_unit_description

```{r}
# create emtpty database structure for raw codes
eutradeflows::createdbstructure(sqlfile = 'raw_comext.sql', dbname = 'tradeflows')
# create empty database structure for validated codes
eutradeflows::createdbstructure(sqlfile = 'vld_comext.sql', dbname = 'tradeflows')
```

## Load table dumps into the database structure

```{r}
if(FALSE){
    path <- '/tmp/comext_sql_dump/'
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_monthly.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_monthly_template.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_partner.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_product.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_reporter.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_unit_description.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_unit.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_comext_yearly_template.sql.7z'))
    eutradeflows::loadtabledump('tradeflows', file.path(path,'raw_flow_yearly.sql.7z'))
}
```

## harvest metadata
The issue was that the metadata was empty for the countries.
There was metadata for products however. 

```{r}
if(FALSE){
    data_raw_path <- "/tmp/data-raw"
    # Connect to the database
    con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
    eutradeflows::createdbstructure(sqlfile = 'raw_comext.sql', dbname = 'tradeflows') 
    tradeharveseter::harvestcomextmetadata(con, data_raw_path)

    # Issues with partner code lead me to load them separately not needed if everything works well
    # Write partner codes to the database
    # tradeharvester::writetxtcodes2db(con, "raw_comext_partner",
    #                                  file.path(data_raw_path, 
    #                                            "/comext/COMEXT_METADATA/CLASSIFICATIONS_AND_RELATIONS/ENGLISH/PARTNERS.txt"),
    #                                  columnnames = c("partnercode", "datestart", "dateend",
    #                                                  "partner", "datestart2", "dateend2"))
    # Disconnect from the database
    RMariaDB::dbDisconnect(con)
}
```

## Clean comext codes

```{r}
if(FALSE){
    # Conect to the DB
    con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
    # Extract unique match between each code and their most recent description
    # Note: the raw data contains repeated codes because the description changes over time. 
    eutradeflows::cleanallcomextcodes(con)
    # Check for the presence of codes 
    reporter <- tbl(con, "vld_comext_reporter") %>% collect()    
    partner <- tbl(con, 'vld_comext_partner') %>% collect()
    product <- tbl(con, 'vld_comext_product') %>% collect()
    unit <- tbl(con, 'vld_comext_unit') %>% collect()
    unit_description <- tbl(con, 'vld_comext_unit_description') %>% collect()
    # Disconnect from the DB 
    RMariaDB::dbDisconnect(con)
}
```


## Clean data
See the other vignette on the [cleaning procedure](clean.html).
