---
title: "dygraph"
author: "Paul Rougieux"
date: "14 February 2018"
output: 
 html_document:
    fig_width: 10
    fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dygraphs)
library(dplyr)
library(eutradeflows)
```


# Load data for IMM product "Sawn: tropical hardwood"

```{r}
productimm_ <- "Sawn: tropical hardwood"
productselected <- classificationimm %>% filter(productimm %in% productimm_)
# Convert to character because SQL is slower if 
# search is performed on character index, while using a numeric value
productselected <- as.character(productselected$productcode)
tableread <- "vld_comext_monthly"
periodstart <- 200000
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
# swd <- eutradeflows::loadflows(con, 440799, periodstart = 200000)
swd <- tbl(con, tableread) %>%
    filter(productcode %in% productselected &
               period > periodstart) %>% 
    addproreppar2tbl(con,.) %>% 
    collect() 
message("Remove trailing white space in country name. Note: this should be fixed in the cleaning procedure")
# Base on https://stackoverflow.com/a/2261349
swd <- swd %>% 
    mutate(reporter = gsub("^\\s+|\\s+$","",reporter),
           partner = gsub("^\\s+|\\s+$","",partner))
# write.csv(filter(swd, reporter=="Utd. Kingdom"), "/tmp/swd.csv")
```
There are `r nrow(swd)` rows in the dataset.

```{r eval=FALSE, include=FALSE}
# Other query
if(FALSE){
    dtf <- tbl(con, "vld_comext_monthly") %>% 
        filter(productcode %in% c("44079927", "44079990", "44079991",
                                  "44079996", "44079998")  & flowcode == 1 ) %>% 
        addproreppar2tbl(con, .) %>% 
        collect() 
productselected <- as.character(productselected$productcode)
tbl(con, tableread) %>%
    filter(productcode %in% productselected &
               period > periodstart) %>% 
    addproreppar2tbl(con,.) %>% 
    explain()
}

```




# Synchronised plots displaying value, quantity and weight 
Uses the [dygraph package](https://rstudio.github.io/dygraphs/index.html).
To zoom, select directly on the plot or move the sliders of the selectin window. 
Double click to zoom out.

The following actions are performed before plotting:
Filter data to keep only imports from Indonesia into the UK.
Aggregate Trade value, quantity and volume (separate data at the 8 digit level is available in the database).
Transform the data frame into a time series "xts" object.
```{r}
# This will be part of the datasetfiltered() reactive function
swd1 <- swd %>% 
    filter(reporter == "Utd. Kingdom" & partner== "Cameroon" & flowcode==1) %>% 
    group_by(reportercode, partnercode, flowcode, unitcode, period) %>% 
    summarise_at(c("tradevalue", "quantity","weight"), sum, na.rm = TRUE) %>% 
    ungroup() %>% 
    mutate(date = lubridate::parse_date_time(period, "ym")) %>% 
    select(date, tradevalue, quantity, weight) %>%
    data.frame()
# Convert to a time series
swd1xts <- xts::xts(swd1[,-1], order.by=swd1[,1])

dateWindow <- c("2012-01-01", "2018-01-01")
dygraph(swd1xts[,"quantity"], main = "Quantity", ylab = "M3", group="tf") %>% 
    dyRangeSelector(dateWindow=dateWindow) %>% 
    dyRoller(rollPeriod=12)

dygraph(swd1xts[,"tradevalue"], main = "Trade Value", ylab = "1000 €", group="tf") %>% 
    dyRangeSelector(dateWindow=dateWindow) %>% 
    dyRoller(rollPeriod=12)

dygraph(swd1xts[,"weight"], main = "Weight", ylab = "T", group="tf") %>% 
    dyRangeSelector(dateWindow=dateWindow) %>% 
    dyRoller(rollPeriod=12)

```


Products included under the IMM name "`r productimm_`"
```{r}
swd %>% 
    filter(reporter == "Utd. Kingdom" & partner== "Cameroon" & flowcode==1) %>% 
    ungroup() %>%  distinct(productcode, productdescription) %>% 
    knitr::kable()
```




# Chips

```{r}
productimm_ <- "Chips"
productselected <- classificationimm %>% filter(productimm %in% productimm_)
# Convert to character because SQL is slower if 
# search is performed on character index, while using a numeric value
productselected <- as.character(productselected$productcode)
tableread <- "vld_comext_monthly"
periodstart <- 200000
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
# swd <- eutradeflows::loadflows(con, 440799, periodstart = 200000)
swd <- tbl(con, tableread) %>%
    filter(productcode %in% productselected &
               period > periodstart) %>% 
    addproreppar2tbl(con,.) %>% 
    collect() 
message("Remove trailing white space in country name. Note: this should be fixed in the cleaning procedure")
# Base on https://stackoverflow.com/a/2261349
swd <- swd %>% 
    mutate(reporter = gsub("^\\s+|\\s+$","",reporter),
           partner = gsub("^\\s+|\\s+$","",partner))
# write.csv(filter(swd, reporter=="Utd. Kingdom"), "/tmp/swd.csv")

```


```{r eval=FALSE, include=FALSE}
# Quantity plot only
swd1 <- swd %>% 
    filter(reporter == "Utd. Kingdom" & partner== "Finland     " & flowcode==1) %>% 
    mutate(date = lubridate::parse_date_time(period, "ym")) %>% 
    select(date, quantity) %>% 
    data.frame()
# Convert to a time series
swd1xts <- xts::xts(swd1[,-1], order.by=swd1[,1])

dygraph(swd1xts, main = "Forest products", ylab = "volume") %>% 
    dyRangeSelector()
```
