---
title: "Price distribution"
output:
  pdf_document: 
    fig_caption: yes
    toc: yes
    toc_depth: 4
---

```{r generate_this_document, eval=FALSE}
# Generate this document from a R command line
rmarkdown::render("~/R/eutradeflows/docs/pricedistribution.Rmd")
#
# Issues ####
message("Issues:")
message("Add price bounds to the plots")
message("Generate density plots which include the price weighted by trade value,
        instead of by number of flows, according to
        https://stackoverflow.com/a/12625991/2641825
        it's possible to pass a `weights` aesthetic to geom_density()
        Acording to 
        https://stats.stackexchange.com/a/22245/68318
        The vector of weights should sum to one for the density to be a true density.")
```


```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir="..") # file paths are relative to the root of the project directory
opts_chunk$set(fig.width = 10, message=FALSE, warning=FALSE)
library(tradeflows)
library(dplyr)
library(tidyr)
library(ggplot2)
``` 

Connect to the database.
```{r dbConnect}
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
```

# Introduction

This document displays price distribution for yearly and monthly Comext data.
Various plots explore the variation of the price distribution among products and through time.
The data is extracted from the database and stored into 2 data frames:

 * wmp contain monthly archive data
 * wpy contain yearly archive data
 
The first plots, Figure 1 and 2 are histograms.
For a given price range, the height of the bar represents the number of flows occuring at that price range. 

Figure 3 and onwards are density plots. 
For a given price range, the area under the curve represents the probability that a given flow has a price withing the range.
Because most of the density is less than 1, 
the curve has to rise higher than 1 in some points 
in order to have a total area of 1 as required for all probability distributions.

# Compare yearly and monthly price distributions for sub products of 4407
```{r}
products <- tbl(con, "vld_comext_product") %>% 
    filter(productcode %like% "4407%")  %>%
    # explain() %>% 
    collect()

# Monthly archive 
wpm <- tbl(con, "raw_comext_monthly_2016S1") %>% 
    filter(productcode %in% products$productcode) %>% 
    collect()

# Monthly recent

# Yearly archive
wpy <- tbl(con, "raw_comext_yearly_2016S2") %>% 
    filter(productcode %in% products$productcode) %>% 
    collect()

# Add prices
wpm <- wpm %>% 
    tradeflows::addconversionfactorandprice()
wpy <- wpy %>% 
    tradeflows::addconversionfactorandprice()

# Bind the monthly and yearly dataframes together
# Specify monthly and yearly timeframe
wpm$timeframe <- "monthly"
wpy$timeframe <- "yearly"
wp <- rbind(wpm, wpy) %>% 
    mutate(year = substr(period, 1, 4))
```

## Descriptive statistics (min, max, quartiles)
Number of rows and total trade value
```{r}
# Number of rows in millions
nrow(wp)/1e6
# Total trade value in million €
sum(wp$tradevalue, na.rm=TRUE) /1e6
# Time frame
unique(wp$year)
# Trade value per year in million €
wp %>% group_by(year, flowcode, timeframe) %>% 
    summarise(tradevalue = round(sum(tradevalue, na.rm=TRUE)/1e6)) %>% 
    left_join(data_frame(flow = c("import", "export"), flowcode = 1:2)) %>% 
    unite(time_flow, timeframe, flow) %>% 
    ungroup() %>% select(-flowcode) %>% 
    spread(time_flow, tradevalue)
```


Descriptive statistics of tradevalue, quantity and weight 
for monthly and yearly data:
```{r}
summary(wpm[c("tradevalue", "quantity", "weight")])
summary(wpy[c("tradevalue", "quantity", "weight")])
```
Descriptive statistics of prices and convertion factor
for monthly and yearly data:
```{r}
summary(wpm[c("price", "pricew", "conversion")])
summary(wpy[c("price", "pricew", "conversion")])
```

\pagebreak

## Price distributions

```{r fig.cap="Price histogram for 0 < price < 10"}
# Interleaved histograms
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_histogram(binwidth=.1, position="dodge") +
    xlim(c(0,10))

ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_histogram(binwidth=.01, position="dodge") +
    xlim(c(0,1))
```



### Density plots of the price variable

```{r fig.cap="Density plot of the price variable"}
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(0,10))

# Same plot scaled between 0 and 1
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(0,1))
```

### Density plot of the price variable weighted by the tradevalue

```{r fig.cap="Density plot of the price variable weighted by the tradevalue"}
ggplot(filter(wp, !is.na(tradevalue)),
       aes(x=price, fill=timeframe, weight = tradevalue/sum(tradevalue))) +
    geom_density(alpha=.3) +
    scale_x_continuous(breaks = c(0.5,0:10), limits = c(0,10))

# Same plot scaled between 0 and 1
ggplot(filter(wp, !is.na(tradevalue)),
       aes(x=price, fill=timeframe, weight = tradevalue/sum(tradevalue))) +
    geom_density(alpha=.3) +
    scale_x_continuous(breaks = 0:10/10, limits = c(0,1))
```


### log(price) distribution monthly and yearly
```{r fig.cap="log(price) distribution monthly and yearly"}
ggplot(wpm, aes(x = log(price))) + 
    geom_histogram(binwidth = 0.1) 

ggplot(wp, aes(x = log(price), fill = timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(-10,5))
```


### Price distribution by product
```{r fig.cap="Price distribution by product, at the 6 digit level"}
wp$productcode6d <- substr(wp$productcode, 1,6)
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(0,1)) +
    facet_wrap(~productcode6d, scales = "free_y")
```

Same plot weighted by the trade value
```{r fig.cap="Monthly distribution  by product, 6 digit level, weighted by trade value"}
wp %>% 
    filter(!is.na(tradevalue)) %>% 
    group_by(productcode6d) %>% 
    mutate(sumtradevalue = sum(tradevalue)) %>%
    # summarise(sum(tradevalue)/unique(sumtradevalue)) # are they all 1?
    ggplot(aes(x=price, fill=timeframe, weight = tradevalue/sumtradevalue)) +
    geom_density(alpha=.3) +
    xlim(c(0,1)) +
    facet_wrap(~productcode6d, scales = "free_y")
```



#### Monthly distribution  by product, 8 digit level
Plot generated externally
Note: some products only have yearly data, no monthly data.
Is this due to missing quantity information? 
Or are the monthly flows missing completely for these products?
```{r eval=FALSE}
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(0,1)) +
    facet_wrap(~productcode, scales = "free_y")
```

Yearly distribution by product

```{r}
# Monthly distribution by year by product
# Yearly distribution by year by product
```

### Price distribution by year


Large plot generated externaly
```{r eval=FALSE}
wp$productcode6d <- substr(wp$productcode, 1,6)
ggplot(wp, aes(x=price, fill=timeframe)) +
    geom_density(alpha=.3) +
    xlim(c(0,1)) +
    facet_grid(year~productcode6d, scales = "free_y")
```



## Ridge plots 

```{r fig.cap= "Price distribution per year, all products", eval=TRUE}
library(ggridges)
# Create a factor of years
wpm$year <- wpm$period %/% 100
wpm$year <- factor(wpm$year, levels = rev(unique(wpm$year)))
ggplot(wpm, aes(x = price, y = year)) +
       geom_density_ridges(scale = 1, alpha = .5) +
    xlim(c(0,1)) +
    theme_ridges()
```

Add price bounds to this plot. 

```{r fig.cap="Price distribution by product at the 6 digit level", fig.height=13}
# Add product code at the 6 digit level
wpm$productcode6d <- substr(wpm$productcode, 1, 6)
ggplot(wpm, aes(x = price, y = year)) +
       geom_density_ridges(scale = 1, alpha = .5) +
    xlim(c(0,1)) +
    theme_ridges() +
    facet_wrap(~productcode6d)
```

Products under 440799 only

```{r fig.cap="Price distribution at the 8 digit level, for sub-products of 440799", fig.height=13}
wpm %>% 
    filter(productcode6d == "440799") %>% 
    ggplot(aes(x = price, y = year)) +
       geom_density_ridges(scale = 1, alpha = .5) +
    xlim(c(0,1)) +
    theme_ridges() +
    facet_wrap(~productcode)
```

\pagebreak


# Price per weight

```{r fig.cap="Price per unit of weight", fig.height=13}
ggplot(wpm, aes(x = pricew, y = year)) +
       geom_density_ridges(scale = 1, alpha = .5,
                           rel_min_height = 0.01) + # Cut trailing tail
    xlim(c(0,10)) +
    theme_ridges() +
    facet_wrap(~productcode6d)
```


Show red bounds using the 5th percentile and 95the percentile
and blue bounds, using half the first quartile and twice the third quartile

```{r}
b <- quantile(wpm$pricew,c(.25,.75), na.rm = TRUE) 
b
b * c(0.5, 2)
quantile(wpm$pricew,c(.05,.95), na.rm = TRUE)

# Group by products
wpm %>% 
    filter(!is.na(pricew)) %>% 
    group_by(productcode) %>% 
    summarise(q5  = quantile(pricew, 0.05, names = FALSE),
              q25 = quantile(pricew, 0.25, names = FALSE),
              q75 = quantile(pricew, 0.75, names = FALSE),
              q95 = quantile(pricew, 0.95, names = FALSE)) %>% 
    kable()
```

```{r}
# With base R,
# see https://stackoverflow.com/questions/43373568/dplyr-to-iterate-over-all-columns-for-quantile
# sapply(wp[!is.na(wp$pricew),],
#        function(x) quantile(x, c(0.25, 0.5, 0.75)))%>% t
# But this doesn't take care of the grouping
```


Disconnect from the database.
```{r dbDisconnect}
RMySQL::dbDisconnect(con)
```
