---
title: "Sankey Flow Diagram"
author: "Paul Rougieux"
date: "8 December 2017"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# knitr::opts_chunk$set(fig.width = 15, fig.height = 9)
library(networkD3)
library(dplyr)
library(eutradeflows)
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
reporter <- tbl(con, "vld_comext_reporter") %>% collect()
partner <- tbl(con, "vld_comext_partner") %>% collect() 
vpaimplementing <- c(302,306,276,700,268,318)
vpanegotiating <- c(272,322,314,488,424,684,701,680,690)
vpa <- c(vpaimplementing, vpanegotiating)
```




## Sawnwood trade between VPA countries and the EU - January to August 2017 
Products under 4407
```{r}
# Sawnwood tradeflows with VPA countries
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '4407%' & 
               flowcode == 1 & 
               period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()

wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```


## Plywood trade between VPA countries and the EU - January to August 2017 
Products under 4412
```{r}
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '4412%' & 
               flowcode == 1 & 
               period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()

wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```



## Furniture trade between VPA countries and the EU - January to August 2017 
Products under the following intems:

* 940330	Furniture; wooden, for office use
* 940340	Furniture; wooden, for kitchen use
* 940350	Furniture; wooden, for bedroom use
* 940360	Furniture; wooden, other than for office, kitchen or bedroom u

```{r}
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '94033%' | 
               productcode %like%  '94034%' |
               productcode %like%  '94035%' |
               productcode %like%  '94036%') %>% 
    filter(flowcode == 1 & period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()

wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```


# Further work planned

* Sankey diagram for many products
* Loop per month
* Loop per product for the sum(2017) 
