---
title: "Sankey Flow Diagram"
author: "Paul Rougieux"
date: "8 December 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(networkD3)
library(dplyr)
# library(tidyverse)
```


# Notes

## Future work
This sankey diagram can be integrated in a 
It could preload data for all years and have a slider for year. 
If the data is not too large, the slider could 

[SHiny network example](https://github.com/christophergandrud/networkD3-shiny-example/blob/master/app.R)

Adding a [slider input with a play button](http://shiny.rstudio.com/reference/shiny/latest/sliderInput.html)

## See also

* original development of the Sankey diagram in the 
Statistical learning RStudio project folder.
* It's possible to use the Sankey diagram in a shiny interface. 



## Errors
Reporter and partner should have different codes.

Error:
It looks like Source/Target is not zero-indexed.
This is required in JavaScript and so your plot may not render.
The problem is that countries in the nodes data frame are
only indexed by position (javascript tables are zero-indexed)
there is no code variable in the nodes table


# Gerating sankey diagram for one product
```{r}
library(eutradeflows)
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
# Raw Eurostat Comext data
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode ==  44072199 & 
               flowcode == 1 & 
               period == "201702") %>% 
    # head(200) %>% 
    collect()

reporter <- tbl(con, "vld_comext_reporter") %>% collect()
partner <- tbl(con, "vld_comext_partner") %>% collect() 

wood2 <- wood %>%
    # Remove partner 1010 and reporter 0
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    preparesankeynodes(reporter, partner, debugname = TRUE)

wood2 %>% plotsankey(value = "tradevalue", units = "K€")
wood2 %>% plotsankey(value = "quantity", units = "M3")

```






# Sankey diagram for aggregated products
```{r}
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '4407%' & 
               flowcode == 1 & 
               period == "201705") %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()
summary(wood$tradevalue)

# Largest sawnwood tradeflows
wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(tradevalue > 2000) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()


```




## January - August 2017 Sawnwood trade between VPA countries and Europe

Cameroon, Central African Republic, Ghana, Indonesia, Liberia, Republic of the Congo
 
  Negotiating
Côte d'Ivoire, Democratic Republic of the Congo, Gabon, Guyana, Honduras, Laos, Malaysia, Thailand, Vietnam
Contact


```{r}
vpaimplementing <- c(302,306,276,700,268,318)
vpanegotiating <- c(272,322,314,488,424,684,701,680,690)
vpa <- c(vpaimplementing, vpanegotiating)



# Sawnwood tradeflows with VPA countries
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '4407%' & 
               flowcode == 1 & 
               period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()
summary(wood$tradevalue)


wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```


## Plywood trade between VPA countries and the EU
```{r}
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '4412%' & 
               flowcode == 1 & 
               period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()
summary(wood$tradevalue)


wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```



## Furniture trade between VPA countries and the EU
940330	Furniture; wooden, for office use
940340	Furniture; wooden, for kitchen use
940350	Furniture; wooden, for bedroom use
940360	Furniture; wooden, other than for office, kitchen or bedroom u

```{r}
wood <- tbl(con, "vld_comext_monthly") %>% 
    filter(productcode %like%  '94033%' | 
               productcode %like%  '94034%' |
               productcode %like%  '94035%' |
               productcode %like%  '94036%') %>% 
    filter(flowcode == 1 & period >= 201701) %>% 
    group_by(reportercode, partnercode) %>% 
    summarise(tradevalue = sum(tradevalue)) %>% 
    collect()
summary(wood$tradevalue)


wood %>% 
    filter(partnercode != 1010 & partnercode != 1011 & 
               reportercode !=0) %>% 
    filter(partnercode %in% vpa) %>%
    preparesankeynodes(reporter, partner) %>% 
    plotsankey()
```




# Sankey diagram for many products

# Loop per month
```{r}

```


# Loop per product for the sum(2017) data
```{r}

```

