---
title: "Missing_Data"
author: "Paul Rougieux"
date: "18 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(eutradeflows)
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
```


# Introduction

See also the same file in the tradeflows package. 


# Issue 5 
[eutradeflows issue 5](https://github.com/stix-global/eutradeflows/issues/5)
"The number of distinct rows for all columns should be equal to the number of distinct codes in the raw dataset"


```{r}
# Recreate the validated database structure
createdbstructure(sqlfile = 'vld_comext.sql', dbname = 'tradeflows')
message("Cleaning product, reporter and partner codes...")
cleancode(con, "raw_comext_product", "vld_comext_product", productcode)
cleancode(con, "raw_comext_reporter", "vld_comext_reporter", reportercode)
message("The issue is with the partner code cleaning")
cleancode(con, "raw_comext_partner", "vld_comext_partner", partnercode)

message("Cleaning unit codes...")
cleancode(con, "raw_comext_unit_description", "vld_comext_unit_description", unitcode)
```

A break point placed in prorepparcodes.R in the `cleancode()` function
shows that this error is due to the fact that code 0 can represent both 
"no data"" and "not determined":
```
vldcode %>% filter(partnercode == 0)
# A tibble: 2 x 2
# Groups:   partnercode [1]
  partnercode partner       
        <int> <chr>         
1           0 No data       
2           0 Not determined
```
"Not determined" is a slightly better qualifier of the absence of partner,
remove "No data" from the raw table.
If "partner" is in the column name and partnercode==0 is duplicated, 
then remove the "No data" from the partner.
```
if("partner" %in% names(rawcode) & sum(rawcode$partnercode==0)>1){
    rawcode <- rawcode[rawcode$partner!="No data",]
}
```



# modify the add unit function

```{r}
productanalysed <- "44071091"
dtfr <- tbl(con, "raw_comext_monthly") %>%
    filter(productcode == productanalysed) %>%
    collect()
    eutradeflows::addunit2tbl(RMariaDBcon,
                              maintbl = .,
                              tableunit = tableunit)  
```

