---
title: "Missing_Data"
author: "Paul Rougieux"
date: "18 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(eutradeflows)
library(dplyr)
library(tidyr)
library(ggplot2)

con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
```


# Introduction

The purpose of this notebook is to compare yearly flows to monthly flows. 

Plan: 
We select a few products of interest, then compare the sum of the monthly trade flows to the yearly trade flows. 
If there are differences, we also compare flags with respect to this sum. 

Products of interest
```{r}
products_of_interest <- c(44072969, 44072799)


tbl(con, 'raw_comext_product') %>% 
    filter(productcode %in% products_of_interest) %>% 
    select(-datestart2, -dateend2) %>% 
    knitr::kable(format='markdown',
          caption='Changes to the SAPELLI product code')
```




# Load data

## Update DB if needed
Load yearly dump to the database. This should be done only once. 

```{r eval=FALSE}
# loadtabledump('tradeflows', '~/downloads/raw_comext_yearly.sql.7z')
```
## Metadata
```{r}
# Load partner country codes and names
partner_names <- tbl(con, 'vld_comext_partner') %>% collect()
# Load reporter country codes and names
reporter_names <- tbl(con, 'vld_comext_reporter') %>% collect()
```


## Yearly flows
```{r}
# Load from the database
yearly <- tbl(con, "raw_comext_yearly") %>%
    # Filter import flows from VPA countries and product 44
    filter(flowcode == 1 &
               # Query seems faster when productcode is a character variable
               productcode %in% products_of_interest) %>% 
    collect() 
# Prepare the data
yearly <- yearly %>% 
    left_join(partner_names, 'partnercode') %>% 
    left_join(reporter_names, 'reportercode') %>% 
    mutate(year  = period %/% 100)
```



## Monthly flows
For example for 
```{r}
# Load from the database
monthly <- tbl(con, "raw_comext_monthly") %>%
    # Filter import flows from VPA countries and product 44
    filter(flowcode == 1 &
               # Query seems faster when productcode is a character variable
               productcode %in% products_of_interest) %>% 
    collect()
# Prepare the data
monthly <- monthly %>% 
    left_join(partner_names, 'partnercode') %>% 
    left_join(reporter_names, 'reportercode') %>% 
    mutate(year  = period %/% 100)
```



# Compare yearly and monthly


## Aggregate monthly

```{r}
monthly_agg <- monthly %>% 
    group_by(reportercode, reporter, partnercode, partner, year) %>% 
    # Sum and rename variable to identify monthly values later in the comparison
    summarise(tradevalue_m = sum(tradevalue), 
              weight_m = sum(weight),
              quantity_m = sum(quantity))
```

## Join and compare

```{r}
join_index <- c('reportercode', 'reporter', 'partnercode', 'partner', 'year')
comp <- yearly %>% 
    select(c(join_index, 'tradevalue','weight', 'quantity')) %>% 
    full_join(monthly_agg, by=join_index) %>% 
    mutate(tradevalue_diff = tradevalue - tradevalue_m,
           weight_diff = weight - weight_m,
           quantity_diff = quantity - quantity_m) 

comp_agg <- comp %>% 
    group_by(reporter, year) %>% 
    summarise(tradevalue_diff_ratio = sum(tradevalue_diff)/sum(tradevalue),
              weight_diff_ratio = sum(weight_diff)/sum(weight),
              quantity_diff_ratio = sum(quantity_diff)/sum(quantity))
# comp_agg
```


# Conclusion 


## January 2020
I compared the aggregated monthly flows to the yearly flows. 
In general the sum of the tradevalue, weight and quantity for all monthly flows within one year is 
equal to the yearly tradevalue, weight and quantity respectively. 
This is what we expected. 
The values are strictly identical in most countries. 
There are some countries in which I observed differences, mostly below 1% of the reported tradevalue. 
Except for Poland for which we saw a difference of 1.4 % in tradevalue in 2004. 


