---
title: "Country details"
author: "Paul Rougieux"
date: "18 March 2018"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir="../..") # file paths are relative to the root of the project directory
library(eutradeflows)
library(dplyr)
library(tidyr)
library(ggplot2)
```


# Introduction

Given a selection of comodities, the purpose of this document is to select the main partner countries. 

## Load data 

Load data from rds files (Serialization Interface for R Objects). 
See the notebook [bio_imports_update_cache.Rmd](bio_imports_update_cache.Rmd) to update the cached data.

```{r}
# Load metadata
partner_names <- readRDS("data_raw/env_impact/partner_names.rds")
reporter_names <- readRDS("data_raw/env_impact/reporter_names.rds")
product_names <- readRDS("data_raw/env_impact/product_names.rds")

product_names_HS2 <- product_names %>% 
    mutate(productcode2d = substr(productcode,1,2)) %>% 
    filter(productcode == productcode2d)

# Load data from cache
system.time(
    yearly <- readRDS("data_raw/env_impact/extra_eu_bio_imports.rds")
)
# 15 seconds for 63 countries
#   user  system elapsed 
# 14.999   0.382  15.389 

yearly <- yearly %>% 
    mutate(year = period %/% 100,
           productcode2d = substr(productcode, 1,2)) 
```


## Aggregate

```{r}
main_products <- yearly %>% 
    filter(year > max(year) - 5) %>% 
    # Calculate the sum for each CN8 product code 
    group_by(year, productcode2d, productcode) %>% 
    summarise(tradevalue = sum(tradevalue),
              weight = sum(weight),
              tradevalue_b = round(sum(tradevalue)/1e9,1)) %>% 
    # Calculate the percentage of trade within each CN2 product group and each year
    group_by(productcode2d, year) %>% 
    arrange(productcode2d, year, desc(tradevalue))  %>% 
    mutate(sumtradevalue = sum(tradevalue, na.rm = TRUE),
           tradevalue_pct = round(tradevalue / sumtradevalue,2),
           tradevalue_pct_cumul = cumsum(tradevalue_pct)) %>% 
    # Keep only products representing 60% of the total trade value
    filter(tradevalue_pct_cumul < 0.6) %>% 
    left_join(product_names, by="productcode") %>%
    mutate(productdescription = substr(productdescription, 1, 20)) 
```

```{r}
main_partners <- yearly %>% 
    filter(year > max(year) -5 & 
               # Keep only main products
               productcode %in% main_products$productcode) %>% 
    # Calculate the sum for each CN8 product code 
    group_by(year, productcode, partnercode, partneriso) %>% 
    summarise(tradevalue = sum(tradevalue),
              weight = sum(weight),
              tradevalue_b = round(sum(tradevalue)/1e9,1)) %>% 
    # Calculate the percentage of trade between partners within each CN8 product group and each year
    group_by(productcode, year) %>% 
    arrange(productcode, year, desc(tradevalue))  %>% 
    mutate(sumtradevalue = sum(tradevalue, na.rm = TRUE),
           tradevalue_pct = round(tradevalue / sumtradevalue,2),
           tradevalue_pct_cumul = cumsum(tradevalue_pct)) %>% 
    # Keep only partners representing 60% of the total trade value
    filter(tradevalue_pct_cumul < 0.8) %>% 
    left_join(product_names, by="productcode") %>%
    left_join(partner_names, by="partnercode") %>% 
    mutate(productdescription = substr(productdescription, 1, 20)) 
    
```


# Main partners at 8 digit level

List of products which represent at least 60% of total trade under the given CN2 chapter (group). 


```{r results='asis'}
i = 1
for (this_product in unique(main_products$productcode)){
    i = i+1
    if(i==5) break # Short loop for development
    # Sub section title 
    product_description <- product_names$productdescription[product_names$productcode==this_product]
    cat(sprintf('\n\n## %s %s \n\n',this_product, substr(product_description,1,57)))
    cat(product_description, '\n\n')
    # Filter only products under this CN2 code
    main_partners %>% 
        filter(productcode == this_product) %>% 
        ungroup() %>% 
        mutate(year = ifelse(year==lag(year, default=0),"",year)) %>% 
        select(year, partner, tradevalue, tradevalue_pct, weight) %>% 
        knitr::kable() %>% 
        print()
}
```

