---
title: "Validate Comext data, using the eutradeflows package"
author: "Paul Rougieux"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Validate Comext data, using the eutradeflows package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  knitr:::html_vignette:
    toc: yes
---


```{r buildvignette, eval=FALSE, echo=FALSE}
# To build this vignette run 
devtools::build_vignettes()
# Then rebuild the package CTRL+SHIFT+B so that the vignette is 
# integrated to the package documentation
```


```{r setup, warning=FALSE, message=FALSE}
library(knitr)
# Do not evaluate code chunks below, only display code
opts_chunk$set(eval=FALSE) 
library(dplyr)
```


## Create the database structure
Recreating the empty database structure:

* will not erase the main  raw and validated data. 
 Because only templates are created. 
* it will erase all complementary raw tables raw_comext_partner, raw_comext_product, raw_comext_reporter, raw_comext_unit, raw_comext_unit_description
* and it will erase all complementary validated tables vld_comext_partner, vld_comext_price, vld_comext_priceconversion , vld_comext_pricew , vld_comext_product, vld_comext_reporter, vld_comext_unit, vld_comext_unit_description

```{r}
# create emtpty database structure for raw codes
eutradeflows::createdbstructure(sqlfile = 'raw_comext.sql', dbname = 'tradeflows')
# create empty database structure for validated codes
eutradeflows::createdbstructure(sqlfile = 'vld_comext.sql', dbname = 'tradeflows')
```


## Connect to the database
```{r dbConnect}
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
```


## Load raw comext data and text files into the database
```{r}
# transfer raw codes
tradeharvester::transfertxtcodesfolder2db(con, rawdatacomextfolder = "~/R/tradeharvester/data-raw/comext/201707/text/english/")

# transfer raw codes on the server
tradeharvester::transfertxtcodesfolder2db(con, rawdatacomextfolder = "/mnt/sdb/data-raw/comext/201710/text/english/")
```

## Clean raw data
### Prepare product, unit, reporter, partner codes
```{r}
cleanallcomextcodes(con)

# Check the content of codes
# Display the first 6 lines of all validated `vld̀  tables 
vldtables <- grep("vld", RMySQL::dbListTables(con), value = TRUE)
lapply(vldtables, 
       function(x){
           tbl(con, x) %>% head() %>% collect() %>% kable(caption = x)
       })
```


## Clean trade flows data

Cleaning is implemented in the `tradeflows` package see the help of the function 
`tradeflows::cleancomext()`.
```{r}
tradeflows::cleancomext()
```

The cleaning procedure is run each night by a cron job:
```
0 5 * * * paul Rscript -e "library(tradeflows); cleancomext('tradeflows')" >> ~/log/clean$(date +"\%Y\%m\%d").log 2>&1
```

## Disconnect from the database
```{r dbDisconnect}
RMySQL::dbDisconnect(con)
```

