---
title: "harvest"
author: "Paul Rougieux"
date: "2 August 2017"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
library(dplyr)
```



Content copied from original file in the docs/harvest.Rmd document of the 
tradeharvester package. The tradeharvester package has now been merged to this
eutradeflows package. 


# 2020 Update of the yearly data

In 2020 in noticed many missing data in the monthly data. 
I am not sure if missing data is 
I need also the yearly data to be harvested properly. 

Moving down the call stack:
* harvest calls harvestcomextdata
* harvestcomextdata does those things with the yearly data:
    * calls downloadcomextmonthlyrecent which downloads both monthly and yearly files to 
    the "COMEXT_DATA/PRODUCTS/" folder.
    * moves yearly files to "COMEXT_DATA/PRODUCTS_YEARLY/" (that works) 
    * and transfer7zfolder2db on the yearly data (issue: no yearly data in the db).

Details on each function call:
```
harvest(rawdatafolder = '/mnt/sdb/public', 
        dbname = 'tradeflows', 
        startyear = 2000,
        productcodestart = tradeharvester::products2harvest$productcode)
```

## On the server
Trying to execute transfer7zfolder2db on the yearly data. 
Complete each argument then connect to the server and make that call from the R prompt.

```{r eval=FALSE}
library(tradeharvester)
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
# Prepare path to yearly input data
rawdatafolder = '/mnt/sdb/public'
comextfolder = 'comext/COMEXT_DATA/PRODUCTS_YEARLY/'
folderyearly <- file.path(rawdatafolder, comextfolder)
# Transfer data from the yearly 7z files to the database
transfer7zfolder2db(RMariaDBcon = con,
                    rawdatacomextfolder = folderyearly,
                    productcodestart = tradeharvester::products2harvest$productcode,
                    tablename = "raw_comext_yearly",
                    tabletemplate = "raw_comext_monthly_template")
```
The above function call seems to work. It transfers yearly data from the 7zip files to the database. 
I wonder why the table doesn't appear as a dump file? 

```{r}
rawtables <- RMariaDB::dbListTables(con)
rawtables <- rawtables[grepl("^raw", rawtables)]
# Dump the yearly table
tablename <- "raw_comext_yearly"
eutradeflows::dumptable("tradeflows", tablename, dumpfolder = file.path(rawdatafolder, "sqldump/"))
```

The dumptable() command starts with this output :
```
> eutradeflows::dumptable("tradeflows", tablename, dumpfolder = file.path(rawdatafolder, "sqldump/"))
Copying the table raw_comext_yearly to a .sql.7z file with the bash command:
mysqldump tradeflows raw_comext_yearly | 7zr a -siraw_comext_yearly.sql /mnt/sdb/public/sqldump/raw_comext_yearly.sql.7z


7-Zip (a) [64] 16.02 : Copyright (c) 1999-2016 Igor Pavlov : 2016-05-21
p7zip Version 16.02 (locale=en_US.UTF-8,Utf16=on,HugeFiles=on,64 bits,2 CPUs Intel Core Processor (Haswell, no TSX, IBRS) (306C1),ASM,AES-NI)

Creating archive: /mnt/sdb/public/sqldump/raw_comext_yearly.sql.7z

Items to compress: 1
```
Maybe it's because the disk is full? 
Yes the disk is nearly full and there is most of the space taken by /var/lib:
```
/var/lib$ sudo du -hd1 |sort -rh
44G     .
32G     ./mysql
12G     ./docker
265M    ./fail2ban
109M    ./apt
26M     ./dpkg
...
```
Mysql is inflating in size more that it needs to due to the fact that it doesn't rebuild the index. 
I could reinstall mysql to see if it fixes the issue? 
Otherwise I could simply ask for more space in the / drive. 
Or install that server somewhere else. 

Anyway, seems like I have the yearly data now. 

## On the laptop

### Download yearly files
Download yearly files to a temporary folder "/tmp/COMEXT_DATA/PRODUCTS_YEARLY/". 

```{r eval=FALSE}
library(tradeharvester)
library(dplyr)
data_folder <- '/tmp/COMEXT_DATA/PRODUCTS_YEARLY' 
# Create the data folder, warn if it already exists
dir.create(data_folder, recursive = TRUE)
# Get the list of available files
url_comext_data_folder <- "https://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&dir=comext%2FCOMEXT_DATA%2FPRODUCTS"
yearly_files <- scraplistoffilesincomext(url_comext_data_folder) %>% 
    filter(grepl("52.7z",file))
# Download yearly files
load_comext_bulk <- function(url_comext, file_name, data_folder){
    download.file(url_comext,file.path(data_folder, file_name))
}
# Download one file
#load_comext_bulk(yearly_files$url[1], yearly_files$file[1], data_folder)
# Download all files
purrr::map2(yearly_files$url, yearly_files$file, load_comext_bulk, data_folder)
```

### Transfer yearly files to the laptop's DB
Transfer the files to a new table called
"raw_comext_yearly_env_impact" so I can keep the other analysis in parallel. 

Started at 10h26 on Monday 27 January 2020.
Second year started 10 minutes later 10h38, 2003 at 10h51, 2004 at 10h57.
We will need 200 minutes for 20 years = approx 3 hours. 

Started at around 16h00 on Friday 11 September 2020.
Second year started about 10 minutes later around 16h13.
Third year started at 16h20.


```{r eval=FALSE}
# Copy this chunk at an R command prompt.
# It takes a long time to run.
library(tradeharvester)
library(dplyr)
data_folder <- '/tmp/COMEXT_DATA/PRODUCTS_YEARLY' 
# Short list of products used for a test
# product_codes_of_interest <- c("08", "15", "29", "44", "45", "47", "48")
# Longer list format string as a character variable with first codes starting with zeros.
product_codes_of_interest <- sprintf("%02.0f", 1:63)
# Add some forest-related products necessary for the Bioeconomy Social Accounting Matrices project
product_codes_of_interest <- c(product_codes_of_interest, 
                               c("65", "70", "85", "88", "90", "94", "64", "66", "84", "96"))
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
# Prepare path to yearly input data
data_folder <- '/tmp/COMEXT_DATA/PRODUCTS_YEARLY' 
# Transfer data from the yearly 7z files to the database
transfer7zfolder2db(RMariaDBcon = con,
                    rawdatacomextfolder = data_folder,
                    productcodestart = product_codes_of_interest,
                    tablename = "raw_comext_yearly_env_impact",
                    tabletemplate = "raw_comext_monthly_template")
```

### Harvest metadata

```{r}
library(tradeharvester)
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
# Empty metadata tables
eutradeflows::createdbstructure(sqlfile = 'raw_comext.sql', dbname = 'tradeflows')
# I deactivated the filter so that it copies all product codes
harvestcomextmetadata(con, rawdatafolder = "/tmp/", pause = 0)
# Transfer only i.e. Avoid re-downloading everything
# transfertxtcodesfolder2db(RMariaDBcon = con,
#                           rawdatacomextfolder =  "/tmp/comext/COMEXT_METADATA/CLASSIFICATIONS_AND_RELATIONS/ENGLISH/")
# Validate reporter, partner and product names
eutradeflows::cleanallcomextcodes(con)
RMariaDB::dbDisconnect(con)
```

Check that you have the most recent metadata by looking for the most recent
starting date of product change with this SQL query: 

```{sql}
select max(datestart) from raw_comext_product;
```


# 2020 update of the monthly data

As of July 2020, the latest monthly data available was from April 2020. 

## On the server


### Download monthly files

```{r monthly2020onserver, eval=FALSE}
library(tradeharvester)
library(dplyr)
data_folder <- '/mnt/sdb/public/comext/COMEXT_DATA/PRODUCTS'
# Get the list of available files
url_comext_data_folder <- "https://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&dir=comext%2FCOMEXT_DATA%2FPRODUCTS"
monthly_files <- scraplistoffilesincomext(url_comext_data_folder)

# Download monthly files
load_comext_bulk <- function(url_comext, file_name, data_folder){
    download.file(url_comext,file.path(data_folder, file_name))
}
# Download one file
load_comext_bulk(monthly_files$url[1], monthly_files$file[1], data_folder)
# Download all files
purrr::map2(monthly_files$url, monthly_files$file, load_comext_bulk, data_folder)
```

### Transfer monthly files to the DB

```{r eval=FALSE}
# Copy this chunk at an R command prompt.
# It takes a long time to run.
library(tradeharvester)
library(dplyr)
data_folder <- '/mnt/sdb/public/comext/COMEXT_DATA/PRODUCTS'
# List of products used for a test
product_codes_of_interest <- c("44", "94")
con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
# Transfer data from the yearly 7z files to the database
transfer7zfolder2db(RMariaDBcon = con,
                    rawdatacomextfolder = data_folder,
                    productcodestart = product_codes_of_interest,
                    tablename = "raw_comext_monthly",
                    tabletemplate = "raw_comext_monthly_template")
```

Check period as they are loaded

```{sql}
select distinct(period) from raw_comext_monthly;
```


### Dump the DB table

Only for the products of interest.



## On the laptop

There is an issue with download dropping because files are too large, or the
download is too slow on my network. Using a VPN doesn't solve the issue.

```{r monthly2020, eval=FALSE}
library(tradeharvester)
library(dplyr)
data_folder <- '/tmp/COMEXT_DATA/PRODUCTS_YEARLY' 
# Create the data folder, warn if it already exists
dir.create(data_folder, recursive = TRUE)
# Get the list of available files
url_comext_data_folder <- "https://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&dir=comext%2FCOMEXT_DATA%2FPRODUCTS"
monthly_files <- scraplistoffilesincomext(url_comext_data_folder)

# Download monthly files
load_comext_bulk <- function(url_comext, file_name, data_folder){
    download.file(url_comext,file.path(data_folder, file_name))
}
# Download one file
#load_comext_bulk(monthly_files$url[1], monthly_files$file[1], data_folder)
# Download all files
purrr::map2(yearly_files$url, yearly_files$file, load_comext_bulk, data_folder)
```

The download drops after a few Mb.
There area 2 versions of the URL, one with file= and
another with downfile=, not sure if this makes a difference. 

https://ec.europa.eu/eurostat/estat-navtree-portlet-prod/BulkDownloadListing?sort=1&file=comext%2FCOMEXT_DATA%2FPRODUCTS%2Ffull201909.7z

h
