---
output:
  pdf_document: 
    fig_caption: yes
    toc: yes
  html_document: default
---
```{r databasetableanalysed, echo=FALSE}
# Comment out to pass tableanalysed as a parameter in the parent environment
tableanalysed = 'raw_comext_monthly_201708' 
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
```
---
title: "Count rows in the database table `r tableanalysed`"
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir="..") # file paths are relative to the root of the project directory
opts_chunk$set(echo = FALSE)
library(eutradeflows)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
``` 


# Introduction
```{r echo = FALSE}
tan <- tbl(con, tableanalysed) %>% 
    summarise(n = n()) %>% 
    collect() %$% n

```

The database table `r tableanalysed` contains `r format(tan, big.mark = "'")` rows. 
These rows represent trade flows between 
`r dbndistinct(con, tableanalysed, reportercode)` reporter countries 
and `r dbndistinct(con, tableanalysed, partnercode)` partner countries,
for `r dbndistinct(con, tableanalysed, productcode)` products
over a period of `r dbndistinct(con, tableanalysed, period)` months.

\pagebreak

# Number of rows per product
We analyse product codes in 3 separate categories: 

 * 2 digit product codes 
 * Special product codes containing characters 
 * 8 digit product codes in 2 ways:
    * Number of rows shown first summarised at the 4 digit level
    * then detailed for each product code at the 8 digit level 
 
```{r}
tapro <- tbl(con, tableanalysed) %>% 
    count(productcode) %>% 
    collect()

tapro2 <- tapro %>% 
    filter(nchar(productcode) == 2)

taprospecial <- tapro %>%
    filter(grepl("[[:alpha:]]", productcode))

tapro8 <- tapro %>% 
    filter(nchar(productcode) == 8 & 
               !grepl("[[:alpha:]]", productcode)) %>% 
    mutate(prod2 = substr(productcode, 1,2),
           prod4 = substr(productcode, 1,4))

# Check if 3 table combined contain the same number of rows as the original
stopifnot(nrow(tapro) == nrow(tapro2) + nrow(taprospecial) + nrow(tapro8))

tapro2 %>% 
    kable(format.args = list(big.mark = "'"),
          col.names = c("Product Code 2 Digits |", "Number of Rows"))

taprospecial %>% 
    kable(format.args = list(big.mark = "'"),
          col.names = c("Special Product Code |", "Number of Rows"))
```

\pagebreak
```{r}
options(knitr.kable.NA = '')
tapro8 %>% 
    group_by(prod2, prod4) %>% 
    summarise(n = sum(n)) %>% 
    mutate(prod02 = substr(prod4,3,4)) %>% 
    gather(key, value, -prod2,-prod02) %>% 
    mutate(key = factor(key, levels= c("prod4","n"))) %>% 
    reshape2::dcast(prod02 ~ prod2 + key) %>% 
    # tidyr::unite(prod4n, prod4, n) %>% 
    # spread(prod2, prod4n, fill = "") %>% 
    mutate_at(vars(ends_with("n")), as.numeric) %>% 
    select(-prod02) %>%
    kable(format.args = list(decimal.mark = ",", big.mark = "'"),
          col.names = c("44", "nrow",
                        "45", "nrow",
                        "47", "nrow", 
                        "48", "nrow",
                        "94", "nrow"))
```



```{r results='asis'}
taprounited <- tapro8 %>% 
    arrange(desc(n)) %>% 
    unite(prod8n, productcode, n) %>% 
    group_by(prod2) %>% 
    summarise(prod8n = paste(prod8n, collapse = ", ")) 

for (prod2 in taprounited$prod2) {
    cat("\n\n## Number of rows for sub-products of product ",
        prod2, "\n\n")
    cat(taprounited$prod8n[taprounited$prod2 == prod2])
    cat("\n\n")
}
```


# Number of rows per period
```{r}
tbl(con, tableanalysed) %>% 
    count(period) %>% collect() %>% 
    mutate(period = as.character(period)) %>% 
    ggplot(aes(x = period, y = n/1e3)) +
    geom_point() +
    ylim(c(0,NA)) +
    ylab("Number of rows in thousand") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
perpro <- tbl(con, tableanalysed) %>% 
    count(period, productcode) %>% 
    collect() %>%
    ungroup() %>% 
    mutate(period = as.character(period))

perpro2 <- perpro %>%
    filter(nchar(productcode) == 2) %>% 
    mutate(productcode = as.numeric(productcode)) %>% 
    left_join(tradeharvester::products2harvest, by="productcode") %>% 
    unite(product, productcode, productdescription, sep = " ")

perprospecial <- perpro %>%
    filter(grepl("[[:alpha:]]", productcode))

perpro8 <- perpro %>% 
    filter(nchar(productcode) == 8 & 
               !grepl("[[:alpha:]]", productcode)) %>% 
    mutate(prod2 = substr(productcode, 1,2),
           prod4 = substr(productcode, 1,4))
stopifnot(identical(nrow(perpro),
                    nrow(perpro2) + nrow(perprospecial) + nrow(perpro8)))
```


```{r fig.height=10, fig.cap="Number of rows for 2 digit produt codes"}
ggplot(perpro2, aes(x = period, y = n)) +
    geom_point() +
    facet_wrap(~product, ncol = 2) +
    ylim(c(0,NA)) +
    ylab("Number of rows") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


```{r fig.height=10, fig.cap="Number of rows for special product codes"}
ggplot(perprospecial, aes(x = period, y = n)) +
    geom_point() +
    facet_wrap(~productcode) +
    ylim(c(0,NA)) +
    ylab("Number of rows") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


# Number of rows per reporter
```{r}

```



```{r dbDisconnect}
RMySQL::dbDisconnect(con)
```

