---
title: "Conversion factor distribution"
output:
  pdf_document: 
    fig_caption: yes
    toc: yes
    toc_depth: 4
---

```{r generate_this_document, eval=FALSE}
# Generate this document from a R command line
rmarkdown::render("~/R/eutradeflows/docs/cvfdistribution.Rmd")
#
```


```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir="..") # file paths are relative to the root of the project directory
opts_chunk$set(fig.width = 10)
library(tradeflows)
library(eutradeflows)
library(dplyr)
library(tidyr)
library(ggplot2)
``` 

Connect to the database.
```{r dbConnect}
con <- RMySQL::dbConnect(RMySQL::MySQL(), dbname = "tradeflows")
```


# Compare yearly and monthly conversion factor for sub products of 4407
```{r}
products <- tbl(con, "vld_comext_product") %>% 
    filter(productcode %like% "4407%")  %>%
    # explain() %>% 
    collect()

# Monthly archive 
wpm <- tbl(con, "raw_comext_monthly_2016S1") %>% 
    filter(productcode %in% products$productcode) %>% 
    # Add quantity units
    addunit2tbl(con, maintbl = ., tableunit = "vld_comext_unit")  %>%
    collect()

# Yearly archive
wpy <- tbl(con, "raw_comext_yearly_2016S2") %>% 
    filter(productcode %in% products$productcode) %>% 
    # Add quantity units
    addunit2tbl(con, maintbl = ., tableunit = "vld_comext_unit")  %>%
    collect()

# Add prices
wpm <- wpm %>% 
    tradeflows::addconversionfactorandprice()
wpy <- wpy %>% 
    tradeflows::addconversionfactorandprice()

# Bind the monthly and yearly dataframes together
# Specify monthly and yearly timeframe
wpm$timeframe <- "monthly"
wpy$timeframe <- "yearly"
wp <- rbind(wpm, wpy) %>% 
    mutate(year = substr(period, 1, 4))

# 
cvf <- wpm %>% 
    mutate(year = substr(period, 1, 4)) %>% 
    tradeflows::extractconversionfactors(dtf, 
                                         grouping = c("productcode", "flowcode",
                                                      "year", "unit"))
```

## Descriptive statistics (min, max, quartiles)
Number of rows and total trade value per products at the 4 and 8 digit level
More details in the countrows.pdf and productcodes.pdf documents.
```{r}
```


Descriptive statistics of tradevalue, quantity and weight 
for monthly and yearly data:
```{r}
summary(wpm[c("tradevalue", "quantity", "weight")])
summary(wpy[c("tradevalue", "quantity", "weight")])
```
Descriptive statistics of prices and convertion factor
for monthly and yearly data:
```{r}
summary(wpm[c("price", "pricew", "conversion")])
summary(wpy[c("price", "pricew", "conversion")])
```

\pagebreak

## Correlation between weight and quantity for the various products

```{r}
ggplot(wp, aes(x = quantity, y = weight, color = timeframe)) + 
    geom_point()
```


### Separate plots at the 6 digit level
The axis scales of each sub-plot are different. 
The slope of the dotted black line is alayws the same at 0.6, 
corresponding the the median conversion factor. 
The slope of the 

```{r}
wp$productcode6d <- substr(wp$productcode, 1,6)

ggplot(wp, aes(x = quantity, y = weight, color = timeframe)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 0.6) + # median conversion factor
    facet_wrap(~productcode6d, scales = "free")
```

### Only sub products of 440710
```{r}
wp %>% 
    filter(productcode6d == 440710) %>% 
    ggplot(aes(x = quantity, y = weight, color = timeframe)) + 
    geom_point() +
    facet_wrap(~productcode, scale = "free")+
    geom_abline(intercept = 0, slope = 0.6) # median conversion factor
```

### Only sub products of 440729
```{r}
wp %>% 
    filter(productcode6d == 440729) %>% 
    ggplot(aes(x = quantity, y = weight, color = timeframe)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 0.6) + # median conversion factor
    geom_abline(data = intercept = 0, slope = 0.6)
    facet_wrap(~productcode, scale = "free")

```



#### Att the 8 digit level
Large plot available outside this document. 
```{r eval=FALSE}

```


# Disconnect from the database.
```{r dbDisconnect}
RMySQL::dbDisconnect(con)
```
