---
title: "Missing_Data"
author: "Paul Rougieux"
date: "18 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(eutradeflows)
library(dplyr)
library(tidyr)
library(ggplot2)

con <- RMariaDB::dbConnect(RMariaDB::MariaDB(), dbname = "tradeflows")
```


# Introduction

TODO move this file to a new git repository for the work on social accountring matrices.

The purpose of this notebook is to export trade prices at an aggregate level for socio accounting matrices. 


1. forestry
2. wood industry --> issue of residues which are joint production
   One activity and several products: main product and by product.
3. pulp and paper

Product aggregares commonly used in partial equilibrium models of the forest sector. 

## forestproductsdemand

Looking at previous work in the repository forestproductsdemand

In `forestproductsdemand/R/load.R`

```
~/rp/forestproductsdemand$ grep downloadfao R/load.R
    newsprint = downloadfao(1671, 5510, 5610, 5622, 5910, 5922)
    printingAndWritingPaper = downloadfao(1674, 5510, 5610, 5622, 5910, 5922)
    otherPaperAndPaperboard = downloadfao(1675, 5510, 5610, 5622, 5910, 5922)
    totalPaperAndPaperboard = downloadfao(1876, 5510, 5610, 5622, 5910, 5922)
    sawnwoodC = downloadfao(1632, 5516, 5616, 5622, 5916, 5922)
    sawnwoodNC = downloadfao(1633, 5516, 5616, 5622, 5916, 5922)
    sawnwoodTotal = downloadfao(1872, 5516, 5616, 5622, 5916, 5922)
#     roundwoodC = downloadfao(1862, 5516, 5616, 5622, 5916, 5922)
#     roundwoodNC = downloadfao(1863, 5516, 5616, 5622, 5916, 5922)
    fuelwood = downloadfao(1864, 5516, 5616, 5622, 5916, 5922)
    indusroundwoodtotal = downloadfao(1866, 5516, 5616, 5622, 5916, 5922)
    indusroundwoodC = downloadfao(1866, 5516, 5616, 5622, 5916, 5922)
    indusroundwoodNC = downloadfao(1867, 5516, 5616, 5622, 5916, 5922)
    indusroundwoodother = downloadfao(1871, 5516, 5616, 5622, 5916, 5922)
    roundwoodTotal = downloadfao(1861, 5516, 5616, 5622, 5916, 5922)
    fibreboard = downloadfao(1874, 5516, 5616, 5622, 5916, 5922)
    particleboard = downloadfao(1646, 5516, 5616, 5622, 5916, 5922)
    plywood = downloadfao(1640, 5516, 5616, 5622, 5916, 5922)
    veneersheets = downloadfao(1634, 5516, 5616, 5622, 5916, 5922)
    woodbasedpanels = downloadfao(1873, 5516, 5616, 5622, 5916, 5922)
    wastepaper = downloadfao(1669, 5516, 5616, 5622, 5916, 5922)
    # but downloadfao doesn't load production and trade quantity.
    fuelwood = downloadfao(1864, 5516, 5616, 5622, 5916, 5922)
```


In `forestproductsdemand/R/clean.R`
The clean() function prepares forest products data from FAOSTAT:
```
   # Load raw data
    load("data-raw/paperandpaperboard.RData")
    load("data-raw/sawnwood.RData")
    load("data-raw/roundwood.RData")
    load("data-raw/woodpanels.RData")
    load("data-raw/wastepaper.RData")
    load("data-raw/fuelwood.RData")

    # Clean each product
    paperproducts <- cleanrawdata(paperandpaperboard, WB,
                                  metadata = data.frame(unit = "Tons", 
                                                  title = "Paper and Paperboard", 
                                                  baseyear = baseyear))
    roundwood <- cleanrawdata(roundwood, WB,
                              metadata = data.frame(unit = "M3", 
                                              title = "Roundwood", 
                                              baseyear = baseyear))
    sawnwood <- cleanrawdata(sawnwood, WB,
                             metadata = data.frame(unit = "M3", 
                                             title = "Sawnwood", 
                                             baseyear = baseyear))
    woodpanels <- cleanrawdata(woodpanels, WB,
                               metadata = data.frame(unit = "M3", 
                                               title = "Wood Panels", 
                                               baseyear = baseyear))
    fuelwood <- cleanrawdata(fuelwood, WB,
                             metadata = data.frame(unit = "M3",
                                                   title = "Fuel Wood",
                                                   baseyear = baseyear))
```



# Load data
```{r}

```


# Compute prices


